The Environmental Dependence of Structures for Massive Galaxies from the Hyper Suprime-Cam Survey

### 2016-07-23 draft/draft_v0.tex
Start to work on sample selection section

% Song Huang's definition 
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\aa{{A\&A}}
\def\aas{{ A\&AS}}
\def\aj{{AJ}}
\def\al{$\alpha$}
\def\bet{$\beta$}
\def\amin{$^\prime$}
\def\annrev{{ARA\&A}}
\def\apj{{ApJ}}
\def\apjs{{ApJS}}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\cc{cm$^{-3}$}
\def\etal{{\ et al.~}}
\def\flamb{erg s$^{-1}$ cm$^{-2}$ \AA$^{-1}$}
\def\flux{erg s$^{-1}$ cm$^{-2}$}
\def\fnu{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$}
\def\hst{{\it HST}}
\def\kms{km s$^{-1}$}
\def\lamb{$\lambda$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over \sim \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over \sim \;$}}
\def\micron{{$\mu$m}}
\def\mnras{{MNRAS}}
\def\nat{{Nature}}
\def\pasp{{PASP}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\pp{\parshape 2 0truein 6.1truein .3truein 5.5truein}
\def\reference{\noindent\pp}
\def\refindent{\par\noindent\parskip=2pt\hangindent=3pc\hangafter=1 }
\def\sb{mag~arcsec$^{-2}$}
\def\solum{$L_\odot$}
\def\solmass{$M_\odot$}
\def\sigs{$\sigma_*$}
\newcommand{\lt}{<}
\newcommand{\gt}{>}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}
\def\redm{\texttt{redMaPPer}}
\def\redbcg{$\Lambda \ge 30$}
\def\nonbcg{$\Lambda < 30$}

% Commenting:
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO: #1}}}
\newcommand{\plan}[1]{\textcolor{cyan}{#1}}
\newcommand{\addref}[1]{\textcolor{red}{REF}}
\newcommand{\comment}[2]{\textcolor{blue}{\textbf{[Comment (#1): #2]}}}
\newcommand{\song}[1]{\textcolor{magenta}{\textbf{[Song: #1]}}}
\newcommand{\alexie}[1]{\textcolor{blue}{\textbf{[Alexie: #1]}}}
\newcommand{\kevin}[1]{\textcolor{green}{\textbf{[Kevin: #1]}}}

@@

\begin{abstract}
    \todo{Place holder; This is the abstract for talk}\\
    Under the most popular formation scenario, the structures of massive central galaxies
    depend on environment at fixed stellar mass due to different assembly histories shaped
    by their host halos. Yet, clear evidence of such effect is still lacking. Using deep,
    multi-band images for a large sample of massive galaxies at 0.2 < z < 0.5 from the
    Hyper Suprime-Cam (HSC) survey, we discover subtle, but systematic structural
    difference for massive galaxies in low and high mass haloes. The differences are
    consistent with richer merger history in more massive halos. We show that the average
    profiles of mass, shape, and color, along with relations among masses within different
    radius (as proxy of mass assembled at different time) can help us gain more insights
    of their assembly history, and the weak lensing analysis enabled by HSC survey further
    helps us connect the differences we find to the average halo properties.
\end{abstract}
\keywords{galaxies: elliptical and lenticular, cD --- galaxies: formation --- 
          galaxies: photometry --- galaxies: structure --- galaxies: surveys}

@@

    \plan{Scientific Background}
    \begin{itemize}
        \item \plan{Massive galaxies are important cosmic probe and unique labs to study 
            galaxy evolution.} 
        \item \plan{Briefly explain why massive early-type galaxies are important through 
            the difficulties of stellar-halo mass relation and stellar mass function at the 
            high-mass end.}
        \item \plan{Brief summary of the current understanding of thieir cosmic assembly 
            history.}
        \item \plan{Explain why we should care about the environment, and why we expect to 
            see some environemtnal dependence in the structure and other properties of
            massive galaxies.}
        \item \plan{Brief review of the current observations.  It is still not clear 
            whether there is a clear environmental dependence.}
    \end{itemize}

@@

    The paper is organized as follows. Section~2 gives a brief overview of the HSC
    observation and data reduction.  We will also summarize the process of sample
    selection.  In Section~3, we will describe the method for deriving the stellar mass 
    surface density profile.  The main results are summarized in Section~4.  Section~5 
    provides discussions of the assumptions used in this work, the potentially interesting
    physical implications, and several future improvements, ending with a sumary in
    Section~6.

    Within this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_m=0.3$, and
    ${\Omega}_{\Lambda}=0.7$.

@@

    \todo{Brief introduction, see the same section in other people's work} \\

    This work makes use of the internal data release \texttt{S15B}, which covers $\sim 100$
    square of degree of sky in all 5-band to the required depth of \texttt{WIDE} survey.

    \todo{The following part is borrowed from Chan's work; Need to re-write}\\
    HSC is the new prime-focus camera on the 8.2-m Subaru telescope located at Mauna Kea, 
    Hawaii.  The field of view (FOV) is 1.5 deg in diameter (\addref: Miyazaki\etal 2012) 
    which is ten times larger than the FOV of Suprime-Cam (\addref: Miyazaki\etal 2002). 
    The HSC Subaru Strategic Program (SSP) survey consists of three layers with different 
    depths: wide, deep and ultra-deep. The wide survey will cover $\sim1400$ deg$^2$ to 
    $r\sim26$ mag in {\it grizY} broad bands.

    The data are processed with \texttt{hscPipe 4.0.1}, a derivative of the Large Synoptic 
    Survey Telescope (LSST) pipeline (\addref: Ivezic\etal 2008, Axelrod\etal 2010), 
    modified for use with Suprime-Cam and Hyper Suprime-Cam.  The photometric calibration 
    is based on data obtained from the Panoramic Survey Telescope and Rapid Response 
    System (Pan-STARRS) 1 imaging survey (\addref: Schlafly\etal 2012, Tonry\etal 2012, 
    Magnier\etal 2013).

    The pixel scale of the reduced images in the data releases is 0.168$\arcsec$.

    \plan{\textbf{Fig.1}: Comparison of SDSS and HSC color image}

\subsection{Candidates of massive galaxies}

>>>>

% Song Huang's definition 
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\aa{{A\&A}}
\def\aas{{ A\&AS}}
\def\aj{{AJ}}
\def\al{$\alpha$}
\def\bet{$\beta$}
\def\amin{$^\prime$}
\def\annrev{{ARA\&A}}
\def\apj{{ApJ}}
\def\apjs{{ApJS}}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\cc{cm$^{-3}$}
\def\etal{{\ et al.~}}
\def\flamb{erg s$^{-1}$ cm$^{-2}$ \AA$^{-1}$}
\def\flux{erg s$^{-1}$ cm$^{-2}$}
\def\fnu{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$}
\def\hst{{\it HST}}
\def\kms{km s$^{-1}$}
\def\lamb{$\lambda$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over \sim \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over \sim \;$}}
\def\micron{{$\mu$m}}
\def\mnras{{MNRAS}}
\def\nat{{Nature}}
\def\pasp{{PASP}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\pp{\parshape 2 0truein 6.1truein .3truein 5.5truein}
\def\reference{\noindent\pp}
\def\refindent{\par\noindent\parskip=2pt\hangindent=3pc\hangafter=1 }
\def\sb{mag~arcsec$^{-2}$}
\def\solum{$L_\odot$}
\def\solmass{$M_\odot$}
\def\sigs{$\sigma_*$}
\newcommand{\lt}{<}
\newcommand{\gt}{>}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}
\def\redm{\texttt{redMaPPer\ }}
\def\redbcg{$\Lambda \ge 30$}
\def\nonbcg{$\Lambda < 30$}
\def\logms{$\log (M_{\star}/M_{\odot})$}
\def\logmh{$\log (M_{\mathrm{Halo}}/M_{\odot})$}

% Commenting:
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO: #1}}}
\newcommand{\plan}[1]{\textcolor{cyan}{#1}}
\newcommand{\addref}{{\textcolor{red}{REF}}}
\newcommand{\comment}[2]{\textcolor{blue}{\textbf{[Comment (#1): #2]}}}
\newcommand{\song}[1]{\textcolor{magenta}{\textbf{[Song: #1]}}}
\newcommand{\alexie}[1]{\textcolor{blue}{\textbf{[Alexie: #1]}}}
\newcommand{\kevin}[1]{\textcolor{green}{\textbf{[Kevin: #1]}}}

@@

\begin{abstract}
    \todo{Place holder; This is the abstract used for giving a talk.}\\
    Under the most popular formation scenario, the structures of massive central galaxies
    depend on environment at fixed stellar mass due to different assembly histories shaped
    by their host halos. Yet, clear evidence of such effect is still lacking. Using deep,
    multi-band images for a large sample of massive galaxies at 0.2 < z < 0.5 from the
    Hyper Suprime-Cam (HSC) survey, we discover subtle, but systematic structural
    difference for massive galaxies in low and high mass haloes. The differences are
    consistent with richer merger history in more massive halos. We show that the average
    profiles of mass, shape, and color, along with relations among masses within different
    radius (as proxy of mass assembled at different time) can help us gain more insights
    of their assembly history, and the weak lensing analysis enabled by HSC survey further
    helps us connect the differences we find to the average halo properties.
\end{abstract}
\keywords{galaxies: elliptical and lenticular, cD --- galaxies: formation --- 
          galaxies: photometry --- galaxies: structure --- galaxies: surveys}

@@

    \plan{Scientific Background}
    \begin{itemize}
        \item \plan{Massive galaxies are important cosmic probe and unique labs to study 
            galaxy evolution.} 
        \item \plan{Briefly explain why massive early-type galaxies are important through 
            the difficulties of stellar-halo mass relation and stellar mass function at the 
            high-mass end.}
        \item \plan{Brief summary of the current understanding of their cosmic assembly 
            history.}
        \item \plan{Explain why we should care about the environment, and why we expect to 
            see some environemtnal dependence in the structure and other properties of
            massive galaxies.}
        \item \plan{Brief review of the current observations.  It is still not clear 
            whether there is a clear environmental dependence.}
    \end{itemize}

@@

    The paper is organized as follows. Section~2 gives a brief overview of the HSC
    observation and data reduction.  We will also summarize the process of sample
    selection.  In Section~3, we will describe the method for deriving the stellar mass 
    surface density profile.  The main results are summarized in Section~4.  Section~5 
    provides discussions of the assumptions used in this work, the potentially interesting
    physical implications, and several future improvements, ending with a summary in
    Section~6.

    All the magnitudes used here are in AB system (\addref).  
    Within this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_m=0.3$, 
    and ${\Omega}_{\Lambda}=0.7$.

@@

    The Subaru Strategic Program (SSP, \addref ) makes use of the new prime-focus 
    camera, the Hyper Suprime-Cam (HSC; \addref: Miyazaki\etal 2012), on the 8.2-m 
    Subaru telescope at Mauna Kea.  
    Taking advantage of the large field of view (FoV; 1.5 deg in diameter) of HSC, 
    this ambitious multi-layer photometric survey will cover $\sim 1400$ deg$^2$ of 
    sky in 5 broad bands (\textit{g r i z Y}) to the depth of $r \sim 26$ mag in the 
    \texttt{WIDE} part in the next few years.  
    This work is based on the internal data release \texttt{S15B}, which covers 
    $\sim 100$ square of degree of sky in all 5-band to the required depth of 
    \texttt{WIDE} survey.  
    The regions covered by this release are overlapped with several previous 
    spectroscopic survey (e.g. SDSS/BOSS \addref; GAMA \addref).

    The data are processed with \texttt{hscPipe 4.0.1}, a derivative of the Large 
    Synoptic Survey Telescope (LSST) pipeline (\addref: Ivezic\etal 2008,
    Axelrod\etal 2010), modified for use with Suprime-Cam and Hyper Suprime-Cam.  
    \texttt{hscPipe} first bias subtract, flat field, model background, and 
    perform object detection and measurement on the single exposure data. 
    Then, different exposures are warped onto a common World Coordinate System 
    (WCS) and combined into final images with improved signal-to-noise ratio 
    (SNR) after astrometric and photometric calibration.  
    The pixel scale of the combined images is 0.168$\arcsec$.
    The photometric calibration is based on data obtained from the Panoramic 
    Survey Telescope and Rapid Response System (Pan-STARRS) 1 imaging survey 
    (\addref: Schlafly\etal 2012, Tonry\etal 2012, Magnier\etal 2013). 
    To achieve consistent deblending and photometry across all bands, the 
    \texttt{hscPipe} will perform multi-band post-processing on the combined 
    images.  
    The footprints and peaks of detected sources on each band will be 
    merged into a single catalog.    
    This consistent set of peaks and footprints is used as starting point 
    for deblend and measure objects on the combined images of each band.  
    These measurements are then merged into a reference catalog. 
    After fixing the centroids, shape, and other non-amplitude parameters 
    of every object in this catalog, \texttt{hscPipe} will derive forced 
    photometry measurements at each band. 
    Details of \texttt{hscPipe} and the multi-band processing method will 
    be presented in \addref (Bosch\etal 2017?).

    In Figure 1, we compare the false color ({\tt gri}) images of three nearby 
    massive ETGs from both SDSS and HSC surveys.  
    The images are generated using similar scaling and color schemes, and clearly 
    demonstrate HSC's capability to reach to low surface brightness domain.  
    In $i$-band (ignoring the slight difference in filter response curve), the 
    HSC \texttt{WIDE} image is $2.5$-$3.0$ magnitude deeper than SDSS on average.  
    This gives us huge advantage in exploring the outskirt of massive ETGs given 
    their extended, shallow outer surface brightness profiles.

    Motivated by the requirement of weak lensing analysis, the $i$-band data 
    typically has the best seeing in all five bands (the median seeing is around 
    FWHM$\sim 0.8\arcsec$).  Therefore, we will mostly use the $i$-band images
    to study the structure of galaxies.

\subsection{Massive central galaxies in different environments}

    The main scientific goal of this work is to investigate mass assembly history 
    of massive, central ETGs, and its dependence on environments (or host dark halo 
    mass) through their the spatial distribution of stellar mass.

    To achieve this goal, we will select \logms$ > 11.5$ massive central 
    galaxies within host halo mass larger and smaller than \logmh$= 14.0$
    at $0.2 < z < 0.5$.  
    As both $M_{\star}$-$M_{\mathrm{Halo}}$ relation and stellar mass function 
    are still quite uncertain at \logms$ > 11.5$, it is of great interest to 
    investigate the structure of galaxies in this region carefully.
    And, this stellar mass cut can greatly reduce the contamination of 
    satellites (e.g. \addref: van~Uitert\etal 2016).  
    Meanwhile, under the adopted cosmology, $1\arcsec$ equals 3.3 and 6.1 kpc 
    at redshit 0.2 and 0.5.  
    Hence, our redshift bin enables us to reliable measure the total stellar 
    mass within the inner 5 to 10 kpc (where the ``in-situ'' component should 
    still dominates) of massive galaxies at the high redshift end assuming the
    typical seeing of data. 
    At the same time, the depth of the data still allows to study the very 
    outskirt ($\sim 100$ kpc) of these galaxies at $z\sim 0.5$.  
    Also, such limits on mass and redshift ensure us a sizable sample while let 
    us safely ignore any significant structural evolution (no star formation, 
    lower merger rate \etal. e.g. \addref).

    Based on recent constraints of $M_{\star}$-$M_{\mathrm{Halo}}$ 
    relation (e.g. \addref), above \logms$ = 11.5$, there is a large scatter 
    of halo mass at fixed stellar mass.  
    At \logmh$<14.0$, these massive galaxies are mostly the centrals of small 
    groups; while at \logmh$>14.0$, they start to become the centrals of very 
    massive groups and galaxy clusters.   
    Limited by the sample size and the capability of measureing halo mass for 
    individual central galaxy, we will simply separate the sample into two 
    broad halo mass bins with the help of the \redm cluster catalog 
    \footnote{See: http://risa.stanford.edu/redmapper/} (
    \texttt{v5.10}, e.g. \addref: Rykoff\etal 2014; Rozo\etal 2015b).
    These clusters are selected from the SDSS DR8 photometric data 
    using overdensity of red-sequence galaxies. 
    For each cluster, the catalog provides robust estimations of 
    photometric redshift $z_{\Lambda}$ and richness $\Lambda$, along with 
    the best candidate of the central galaxy (the one with the highest 
    central probability \texttt{P\_CEN}).
    Information about the identified member candidates is also provided 
    separately. 
    Please see Rozo\etal (2014, 2015a, 2015b) for more details about the 
    performance of the \redm cluster catalog.      
    Several works (e.g. Li\etal 2015; Saro\etal 2015; Farahi\etal 2016; 
    Simet\etal 2016) have tried to calibrate the $M_{200c}-\Lambda$ relation 
    using different methods.
    Despite the slightly different calibrations derived, it is safe to 
    assume that most clusters identified by \redm ($\Lambda > 20$) have
    $\log (M_{200c}/M_{\odot}) \geq 14.0$.  
    Considering the typical uncertainty of richness, and the fact that 
    \redm catalog starts to become incomplete toward low richness 
    ($\Lambda < 40$) end at $z > 0.33$, we will focus on the $\Lambda > 30$
    clusters in this work.  
    Based on the calibrations from Farahi\etal (2016) or Simet\etal (2016), 
    such richness cut gives us haloes with 
    $M_{200c} \geq 1.56\pm0.35 \times 10^{14} M_{\odot}$ or 
    $M_{200c} \geq 1.60\pm0.11 \times 10^{14} M_{\odot}$.  
    Therefore, \redm can provide us a sample of massive central galaxies in
    \logmh$\geq 14.0$ haloes.  
    And, we assume that, in the same footprints and redshift range, the 
    \logms$\geq 11.5$ galaxies outside \redm clusters are most likely to be 
    the central of \logmh$< 14.0$ haloes.

    According to Leauthaud\etal (2016), all $0.2 < z < 0.5$ and \logms$\geq 11.5$
    galaxies should have $i_{\mathrm{SDSS, cModel}} \leq 21.0$ mag. 
    Ignoring the tiny difference in response curves between SDSS and HSC 
    $i$-band filters, we started by selecting all galaxies with 
    $i_{\mathrm{HSC, cModel}} \leq 21.5$ in the regions that are covered 
    in all five filters, and have already reached the expected depth of 
    \texttt{WIDE} survey.  
    The details of the HSC \texttt{cModel} photometry, and its performence
    will be described in Bosch\etal in. prep. 
    It is quite similar to the SDSS \texttt{cModel} in principle as it also 
    fits the total flux of an object using a combination of de~Vaucouleur 
    and exponential components after the PSF convolution is considered.  
    In our selection, we require well defined centroids and \texttt{cModel} 
    magnitude in all five bands. 
    \song{Should we discuss the impact of failed cModel photometry?}
    A series of quality control cuts are applied to remove objects that are 
    affected by saturation, cosmic ray, other optical artefacts.  
    After these cuts, a total number of \plan{N1} galaxies left in the 
    sample () \todo{Finish this part}  
    (This sample will be referred as \texttt{phoAll}).

    We match this sample with the central galaxies of \redm catalogs using 
    a $2.0\arcsec$ radius, and it results in \plan{N2} galaxies.  
    A small fraction of \redm centrals in the our footprints do not have 
    matched object in our sample due to severe contamination from optical 
    artefacts or bleeding trails from saturated objects.  
    And slight change of the mathcing radius has no impact of this sample.  
    For galaxy in this sample (will be referred as \texttt{redBCG}), 
    the photometric redshift of the cluster is available for each of them; 
    while for \plan of them, the spectroscopic redshift from SDSS DR8 is 
    also provided by \redm catalog when available.

    For the vast majority of galaxies that are not the centrals of very massive
    haloes, we need other sources for their redshfit estimations. 
    Although there has been many efforts for deriving photometric redshift 
    (photo-$z$) using HSC five-band photometry, it is still a working progress 
    now, and its performence at $z \leq 0.5$ is expected to be less satisfying 
    due to the lack of information at the shorter wavelength side of the 
    4000$\AA$ break.  
    Therefore, we will only rely on external spectroscopic redshifts (spec-$z$)
    for this work.  
    The HSC database compiles a catalog of external spec-$z$ for this data 
    release by matching the detected objects with public data of several 
    spectroscopic surveys (e.g. SDSS/BOSS \addref; GAMA \addref; PRIMUS \addref;
    \etal). 
    Duplicated matches from different sources are merged through interal 
    matching using $0.5\arcsec$ radius.

    We match the \texttt{phoAll} sample with this external spec-$z$ catalog 
    using a $1.0 \arcsec$ radius, and it leads to \plan{N3} results. 
    There are \plan{N4} objects at $0.2 \leq z \leq 0.5$.   
    The \texttt{redBCG} sample is also matched in the same way. 
    For ...
    \todo{Finish it...}
